Structural-temporal neighbourhoods of posts to characterize conversations in online forums

### 2016-04-30 doc/IEEEtran/paper_IEEEformat.tex
backup

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Alberto Lumbreras \\ Julien Velcin}
\IEEEauthorblockA{Laboratoire ERIC\\ Université de Lyon, France\\
alberto.lumbreras@univ-lyon2.fr\\
julien.velcin@univ-lyon2.fr}
\and
\IEEEauthorblockN{Bertrand Jouve}
\IEEEauthorblockA{FRAMESPA/IMT\\Université de Toulouse 2, France\\
jouve@univ-tlse2.fr}
\and
\IEEEauthorblockN{Marie Guégan}
\IEEEauthorblockA{Technicolor, France\\
marie.guegan@technicolor.com}}

@@

The remaining of the paper is as follows. We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees. Then we introduce our two definitions of structural-temporal neighbourhoods. To illustrate the kind of neighbourhoods that we obtain in a real data, we analyse the neighbourhood census of a Reddit forum. Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods. We close the paper with some suggestions of future research.
%\begin{figure}
%	\centering
	%\includegraphics[width=0.3\textwidth]{tree1}
%	\includegraphics[width=0.3\textwidth]{tree2}
	%\includegraphics[width=0.3\textwidth]{tree3}
%	\caption{Representation of discussion threads with post tree graphs. Vertex represent posts and edges represent replies between posts.}
%	\label{fig:trees}
%\end{figure}

\section{Discussion trees}
% Introduce el objeto y la nomenclatura
We represent a discussion thread by a tree graph $G=(V,E)$ where $V$ is a set of $n$ vertices representing the posts (also called messages or comments) and $E$ is a set of $n-1$ directed edges that say which posts replied to which post. Vertices have two attributes namely time and author. If for two given vertices $v_i,v_j$ there exists an edge  $e=(v_i, v_j) \in E$ then $v_j$ is called the \textit{parent} of $v_i$, denoted as $p(v_i)=v_j$. The \textit{root} of the tree is the only vertex with no parent, and corresponds to the post that starts the discussion. We say that two vertices $v_i,v_j$ are \textit{siblings} if and only if $p(v_i)=p(v_j)$. A vertex $v_i$ is an ancestor of a vertex $v_j$ and $v_j$ is descendant of $v_i$ if and only if $v_i$ is in the path from $v_j$ to the root. A \textit{leaf} is a post with no replies. A \textit{branch} is the path between a leaf and the root. Two vertices $v_i$ and $v_i$ are said to be \textit{neighbours} if either $p(v_i)=v_j$ or $p(v_j)=v_i$ or, in other words, if their distance in the undirected version of $G$ is one.

@@

\section{Structural neighbourhoods}\label{sec:structural}
Extending the classic definition of neighbourhood according to which two vertices are neighbours if the distance between them is one, we start by the following definition of \textit{structural neighbourhood}:

\begin{definition}
Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed by all the vertices that are at distance equal or less than $r$ from post $i$.
\end{definition}
% pros and cons
In the context of discussion threads, this definition has two limitations. First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written. In conversations, time plays an important role, therefore this is an important limitation. Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically similar.

\section{Structural-temporal neighbourhoods}\label{sec:temporal}
We propose two refinements of the structure-based neighbourhood to take into account the order and time in which posts are written. As we will see, the criteria to choose which posts are neighbours of a post $i$  is not only based on the structural distance but also on the temporal distance to $i$. In the order-based definition, a new radius is set to include at most the $n$ closest posts in time.  In the time-based definition, the new radius is set to include all posts (in the structural-neighbourhood) until the first point where the conversation slows down.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{order_neighbourhood}
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Illustration of order-based (top) and time-based (bottom) neighbourhoods. Grey nodes represent the ego post. Nodes with circles are those in the neighbourhood. Numbers indicate the order of the post. The order-based neighbourhood has parameters $r=3$, $n=6$. The time-based neighbourhood has parameter $r=3$. In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
\label{fig:cutpoints}
\end{figure}
\subsection{Order-based}
\begin{definition}
Given a ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ and order $n$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r,n)$ is the induced subgraph from its structural neighbourhood composed by the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^T(r,n)$.  
\end{definition}
An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints}.
This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are not only near to $i$ in the structure but also in time. Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite. The main limitation of this definition is that the we have no \textit{a priori} criteria to choose the proper parameter $n$ other than making it small to capture only the local dynamic of the conversation around the post $i$.

@@

Rather than looking for a fixed time radius, we may instead decide it by looking at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation. If applied either to the timestamps of posts in a branch (vertical sequence) or to the timestamps of the replies to the same post (horizontal sequence), we say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}.}. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic than $i$.
\end{definition}
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.

% Explain algorithm
Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the above definition. Since the breakpoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical breakpoints in the tree. In the case of multiple branches with some common posts, we consider that a common post is a vertical breakpoint if it is a vertical breakpoint in any of the branches. Once we have the breakpoints, we can proceed with the algorithm. First, we extract the structural-neighbourhood. The time-based neighbourhood will be a subset of the later. Then, we look for horizontal and vertical breakpoints, which mark the frontiers of the time-based neighbourhood. There are four possible cases:

\begin{itemize}
\item A \textit{vertical breakpoint in the ancestors}: If the breakpoint is in the path between the post and the root, then the breakpoint started the new local dynamic to which the ego post belongs. Thus, we remove the ancestors of the breakpoint, but not the breakpoint itself.
\item A \textit{vertical breakpoint in the descendants}: If the breakpoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the breakpoint and the breakpoint itself. 
\item A \textit{horizontal breakpoint either in the older siblings or in the ancestors}: If the breakpoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs. Similarly, if the breakpoint is among the ancestors, then every older sibling of the breakpoint belongs to a previous local dynamic. In both cases, we remove the older siblings of the breakpoint but not the breakpoint itself.
\item A \textit{horizontal breakpoint elsewhere}: In any other case, the horizontal breakpoint starts a different local dynamic and therefore we remove its younger siblings and the breakpoint. 
\end{itemize}
Any time a vertex is removed we also remove its descendants. The final neighbourhood will be connected induced subgraph from the structural neighbourhood. If there are no changepoints in the structural neighbourhood, the time-based neighbourhood will be exactly the structural neighbourhood.

@@

\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $g$, vertical breakpoints, horizontal breakpoints, ego post $ego$
\ENSURE Subgraph of $g$ with all vertices in $V(g)$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors(ego) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings(ego) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical breakpoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal breakpoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{(t)}(r) \leftarrow$ delete(dump\_posts $\cup$ descendants(dump\_posts)) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

@@

\begin{itemize}
\item \textit{Red}: ego post.
\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (when the same author participates several times in the discussion \cite{Backstrom2013})
\item \textit{Yellow}: parent of ego post and other posts written by the same author. It allows to identify, for instance, debates between the ego author (red and orange) and the other author (yellow). We have observed this phenomena in our data, and it is often the cause of long chains.
\item \textit{White}: root post. Differentiating the root post from the rest has been proven by previous research on online discussions. For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}). Also, in terms of preferential attachment, root posts usually get more replies than the non-roots (\cite{Gomez2010, Gomez2012}).
\item \textit{Black}: none of the above.
\end{itemize}
Since some posts might correspond to several colours, we perform a sequential assignment of colours given by $\texttt{black, orange, red, yellow, white}$.

@@

\subsection{Pruning}
At this point, we can still find structures the only difference of which is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same colour and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.

\section{Application to Reddit forums}
Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid on October 18th and 19th, 2014. Nowadays, its members use it mainly to share and discuss about political news. The Game of Thrones forum is a casual discussion forum about the Game of Thrones TV series.
The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between April 25th and October 20th, 2014. The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.

\subsection{Neighbourhood extraction}
As a previous step for all further analysis, we extract the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums. We label every structure with a unique integer number, where a structure is a coloured graph. Two neighbours have the same structure, and therefore the same integer number, if their coloured graphs are isomorphic.

Unfortunately, it is not practical to give a number to every theoretically possible structure since, even if we could enumerate all the possible trees up to a given size, a lot of these trees would not be seen in the data. It appears then a much more practical approach to enumerate only those trees that are seen in the data. In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished. We will call this final labelling a \textit{dictionary} of neighbourhoods.

@@

Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth). Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.

% Computational cost
In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood that already appears in the dictionary before concluding that it needs a new entry. In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts. In practice, however, the bigger the number of posts processed, the less likely we go through all the dictionary without finding an isomorphism. Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find an isomorphism among the first neighbours of the dictionary.

@@

\begin{table}
 \begin{center}
   \tabcolsep = 0.8\tabcolsep
   \begin{tabular}{|cc|cc|cc|}
      \hline
      \multicolumn{2}{|c|}{struct-based} & \multicolumn{2}{|c|}{order-based} &  \multicolumn{2}{|c|}{time-based} \\
      %\hline
     	 POD & GOT 						& POD & GOT 		& POD & GOT\\
   \hline \hline
     1 (12.5)& 3(12.5)   &  1 (12.5) & 1 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  2 (12.5) & 2 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\  
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\    
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\ 
     1 (12.5)& 3(12.5)   &  1 (12.5) & 3 (12.5) &  1 (12.5) & 3 (12.5)\\          
   \hline
   \end{tabular}
\label{tab:table}
 \end{center}
 \caption{Frequence}
\end{table}

>>>>

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Alberto Lumbreras \\ Julien Velcin}
\IEEEauthorblockA{Laboratoire ERIC\\ Université de Lyon, France\\
alberto.lumbreras@univ-lyon2.fr\\
julien.velcin@univ-lyon2.fr}
\and
\IEEEauthorblockN{Bertrand Jouve}
\IEEEauthorblockA{FRAMESPA/IMT\\Université de Toulouse, France\\
jouve@univ-tlse2.fr}
\and
\IEEEauthorblockN{Marie Guégan}
\IEEEauthorblockA{Technicolor, France\\
marie.guegan@technicolor.com}}

@@

The remaining of the paper is as follows. We first discuss about the convenience of the classic neighbourhood definition in dynamic graphs such as discussion trees. Then we introduce our two definitions of structural-temporal neighbourhoods. To illustrate the kind of neighbourhoods that we obtain in a real data, we analyse the neighbourhood census of a Reddit forum. Finally, we apply both neighbourhood definitions to detect clusters of users that tend to appear in the same neighbourhoods. We close the paper with some suggestions of future research.

\section{Discussion trees}
% Introduce el objeto y la nomenclatura
We represent a discussion thread by a tree graph $G=(V,E)$ where $V$ is a set of $n$ vertices representing the posts (also called messages or comments) and $E$ is a set of $n-1$ directed edges that say which posts replied to which post. Vertices have two attributes namely time and author. If for two given vertices $v_i,v_j$ there exists an edge  $e=(v_i, v_j) \in E$ then $v_j$ is called the \textit{parent} of $v_i$, denoted as $p(v_i)=v_j$. The \textit{root} of the tree is the only vertex with no parent, and corresponds to the post that starts the discussion. We say that two vertices $v_i,v_k$ are \textit{siblings} if and only if $p(v_i)=p(v_k)$. A vertex $v_l$ is an ancestor of a vertex $v_i$ and $v_i$ is descendant of $v_l$ if and only if $v_l$ is in the path from $v_i$ to the root. A \textit{leaf} is a post with no replies. A \textit{branch} is the path between a leaf and the root. Two vertices $v_i$ and $v_m$ are said to be \textit{neighbours} if either $p(v_i)=v_m$ or $p(v_m)=v_i$. The neighbours of a vertex $i$ are its parent and its children.

@@

\section{Structural neighbourhoods}\label{sec:structural}
Extending the classic definition of neighbourhood, we recall the definition of \textit{structural neighbourhood}\footnote{We coined the term \textit{structural} to differentiate if from the other neighbourhoods discussed in this paper.}:

\begin{definition}
Given a tree graph $G$, the \textit{structure-based neighbourhood} of radius $r$ of post $i$, denoted as $\mathcal{N}_i(r)$, is the induced graph composed by all the vertices that are at distance equal or less than $r$ from post $i$.
\end{definition}
% pros and cons
In the context of discussion threads, this definition has two limitations. First, the decision on whether to include some post in the neighbourhood is only based on the structural distance, and therefore two posts that are at distance $d\leq r$ are considered neighbours of $i$ regardless of the time when they were written although, in conversations, time plays an important role. Another consequence of looking only at the structure is that the number of possible neighbourhoods within a radius $r$ is infinite. This poses a problem when trying to categorize conversations since many conversations, while structurally different, can be considered semantically similar (e.g.: we might not want to put in different categories two structures representing, respectively, a post with 50 replies and another with 40).

\section{Structural-temporal neighbourhoods}\label{sec:temporal}
We propose two refinements of the structure-based neighbourhood to take into account the order and time at which posts are written. In the order-based definition, a new radius is set to include at most the $n$ closest posts in time.  In the time-based definition, the new radius is set to include all posts (in the structural-neighbourhood) until the first point where the conversation slows down.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{order_neighbourhood}
\includegraphics[width=0.45\textwidth]{breakpoints}
\caption{Illustration of order-based (top) and time-based (bottom) neighbourhoods. Grey nodes represent the ego post. Nodes with circles are those in the neighbourhood. Numbers indicate the order of the post. The order-based neighbourhood has parameters $r=3$, $n=6$. The time-based neighbourhood has parameter $r=3$. In the time-based neighbourhood, horizontal changepoints (yellow) and vertical changepoints (orange) represent posts that are temporally far from their predecessors (siblings or parents) and therefore set the limits of the neighbourhood.}
\label{fig:cutpoints}
\end{figure}
\subsection{Order-based}
\begin{definition}
Given an ordered tree graph $G$, the \textit{order-based neighbourhood} of radius $r$ and order $n$ of vertex $i$, denoted as $\mathcal{N}_{i}^O(r,n)$ is the induced subgraph from its structural neighbourhood $\mathcal{N}_{i}(r)$ composed by the $n$ vertices that are closest to $i$ in time and for which there exists a path to $i$ in $\mathcal{N}_{i}^O(r,n)$.  
\end{definition}
An example of order-based neighbourhood is given in Figure~\ref{fig:cutpoints}.
This definition has two advantages over the \textit{structural neighbourhood}. First, the temporal aspect of the conversation is better taken into account since the neighbourhood only includes posts that are not only near to $i$ in the structure but also in time. Second, the size of the neighbourhood has an upper bound of $min(|\mathcal{N}_i(r))|, n)$ thus making the space of possible neighbourhood structures finite. The main limitation of this definition is that the we have no \textit{a priori} criteria to choose the proper parameter $n$ other than making it small to capture only the local dynamic of the conversation around the post $i$.

@@

Rather than looking for a fixed time radius, we may instead decide it by looking at changes in the pace how new posts are added to the thread. In statistical analysis, a \textit{changepoint} in a sequence $x_1,...x_n$ is a point that comes from a different probability distribution than its precedent values. If sequences are timestamps, they are monotonic increasing, therefore the changepoints will correspond to sudden pauses in the conversation. In the following, we will differentiate \textit{horizontal changepoints}, that arise between siblings, from \textit{vertical changepoints}, that arise within a branch. We say that a sequence posts with timestamps $t_1,...,t_n$ belong to the same (vertical or horizontal) \textit{local dynamic} if there is no changepoint $t_i$ in the sequence such that $1 < i \leq n$. For the detection of changepoints we use the PELT algorithm \cite{Killick2012}\footnote{An implementation of this algorithm written by the authors themselves is available in the R library \texttt{changepoint}. We use their \texttt{cpt.meanvar} function.}. Now we can introduce our second definition of neighbourhood.
\begin{definition}
Given a tree graph $G$, the \textit{time-based neighbourhood} of radius $r$ of vertex $i$, denoted as $\mathcal{N}_{i}^T(r)$, is the maximal subgraph of the structural neighbourhood $\mathcal{N}_i(r)$ where all the vertices  belong to the same vertical and horizontal local dynamic as $i$.
\end{definition}
Note that we still have a radius parameter $r$ to guarantee that the resulting structure remains local even if no changepoint is detected near the post $i$.

% Explain algorithm
Algorithm \ref{alg:temporal_neighbourhood} extracts time-based neighbourhoods of a given post according to the above definition. Since the changepoints only depend on the tree and not on the particular post we analyse, we previously detect the horizontal and vertical changepoints in the tree. In the case of multiple branches with some common posts, we consider that a common post is a vertical changepoint if it is a vertical changepoint in any of the branches. Once we have the changepoints, we can proceed with the algorithm. First, we extract the structural-neighbourhood. The time-based neighbourhood will be a subset of the later. Then, we look for horizontal and vertical changepoints, which mark the frontiers of the time-based neighbourhood. There are four possible cases:

\begin{itemize}
\item A \textit{vertical changepoint in the ancestors}: If the changepoint is in the path between the post and the root, then the changepoint started the new local dynamic to which the ego post belongs. Thus, we remove the ancestors of the changepoint, but not the changepoint itself.
\item A \textit{vertical changepoint in the descendants}: If the changepoint is a descendant then it started a new different dynamic and therefore we must remove the descendants of the changepoint and the changepoint itself. 
\item A \textit{horizontal changepoint either in the older siblings or in the ancestors}: If the changepoint is among the older siblings, then it started the horizontal dynamic to which the ego post belongs. Similarly, if the changepoint is among the ancestors, then every older sibling of the changepoint belongs to a previous local dynamic. In both cases, we remove the older siblings of the changepoint but not the changepoint itself.
\item A \textit{horizontal changepoint elsewhere}: In any other case, the horizontal changepoint starts a different local dynamic and therefore we remove its younger siblings and the changepoint. 
\end{itemize}
Any time a vertex is removed we also remove its descendants. $\mathcal{N}_i^T(r)$ is a connected induced subgraph of $\mathcal{N}_i(r)$.  If there are no changepoints in the structural neighbourhood, then $\mathcal{N}_i^T(r) = \mathcal{N}_i(r)$.

@@

\begin{algorithm}[H]
\begin{algorithmic}
\REQUIRE Posts tree $G$, vertical changepoints, horizontal changepoints, ego post $i$, radius $r$
\ENSURE $\mathcal{N}_i^{T}(r)$: Time-based neighbourhood of $i$ at radius $r$
\STATE Compute structural neighbourhood $\mathcal{N}_i(r)$
\STATE ancestors $\leftarrow$  ancestors($i$) in $\mathcal{N}_i(r)$
\STATE older\_siblings $\leftarrow$ older\_siblings($i$) in $\mathcal{N}_i(r)$
\STATE dump $\leftarrow \varnothing$
\FOR{bp $\in$ vertical changepoints}
 \IF{bp $\in$ ancestors}
   \STATE dump $\leftarrow$ dump $\cup$ ancestors(bp)
 \ELSE
   \STATE dump $\leftarrow$ dump $\cup$ descendants(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\FOR{bp $\in$ horizontal changepoints}
  \IF{bp $\in$ (older\_siblings $\cup$ ancestors)}
     \STATE dump $\leftarrow$ dump $\cup$ older\_siblings(bp)
   \ELSE
     \STATE dump $\leftarrow$ dump $\cup$ younger\_siblings(bp) $\cup$ bp
  \ENDIF
\ENDFOR
\STATE $\mathcal{N}_i^{T}(r) \leftarrow$ delete(dump\_posts $\cup$ descendants(dump\_posts)) from $\mathcal{N}_i(r)$
\end{algorithmic}
\caption{Extraction of time-based neighbourhood}
\label{alg:temporal_neighbourhood}
\end{algorithm}

@@

\begin{itemize}
\item \textit{Red}: ego post.
\item \textit{Orange}: other posts written by the author of ego. It allows to identify re-entries (when the same author participates several times in the discussion \cite{Backstrom2013})
\item \textit{Yellow}: parent of ego post and other posts written by the same author. It allows to identify, for instance, debates between the ego author (red and orange) and this other author (yellow). We have observed this phenomena in our data, and it is often the cause of long chains.
\item \textit{White}: root post. Differentiating the root post from the rest has been proven by previous research on online discussions. For instance, some types of users seem to get more replies than others when they initiate a thread (\cite{Himelboim2009,  Lumbreras2013}). Also, in terms of preferential attachment, root posts usually get more replies than the non-roots (\cite{Gomez2010, Gomez2012}).
\item \textit{Black}: none of the above.
\end{itemize}
Since some posts might correspond to several colours, we perform a sequential assignment of colours given by $\texttt{black, orange, red, yellow, white}$.

@@

\subsection{Pruning}
At this point, we can still find structures the only difference of which is that one of them has more leafs hanging from some of the nodes. Thus, we prune every neighbourhood by leaving a maximum of two consecutive siblings of the same colour and where neither of them has any children. The reason of setting the limit to two is that it is the minimum necessary to distinguish between \textit{zero}, \textit{one} and \textit{more than one} consecutive occurrences. In other words, we consider that the difference between one and zero replies to a post is relevant, but that five and six replies do not make any difference worth being represented.  Moreover, this choice encourage smaller structures, which allows the computation of isomorphisms (Section~\ref{sec:extraction}) even in real time.

\section{Application to Reddit forums}

Our data consists of two different forums of Reddit, namely the Podemos dataset \footnote{https://www.reddit.com/r/podemos} and the Game of Thrones dataset\footnote{https://www.reddit.com/r/gameofthrones}. The Podemos forum was conceived in March 2014 as a tool for internal democracy, and forum members used it to debate ideological and organizational principles that were later formalized in their first party congress hold in Madrid on October 18th and 19th, 2014. Nowadays, its members use it mainly to share and discuss about political news. The Game of Thrones forum is a casual discussion forum about the Game of Thrones TV series.
The Podemos dataset contains 75,000 posts corresponding to 4,262 threads written by 9,160 users between April 25th and October 20th, 2014. The Game of Thrones dataset contains 75,000 posts corresponding to 5,862 threads written by 18,907 users.

\subsection{Neighbourhood extraction}\label{sec:extraction}
As a previous step for all further analysis, we extracted the structure-based ($r=2$), the order-based ($r=2, n=4$) and the time-based ($r=2$) neighbourhood around every post for the two forums.

@@

%We identified every type of neighbourhood with a unique integer (label). Two neighbourhoods have the same label if their coloured graphs are isomorphic.

%In order to share a unique labelling of structures among the two forums and the three types of neighbourhoods (which gives 6 different labellings), we re-label the structures after the extraction for every post is finished. We will call this final labelling a \textit{dictionary} of neighbourhoods.

The set of all neighbourhoods $\mathcal{N}_i(r)$, $\mathcal{N}_i^O(r,n)$ and  $\mathcal{N}_i^T(r)$ for the two forums constitutes our \textit{dictionary}.

Once the six extractions are finished, we merge the dictionaries into a global one and sort the labels by frequency of the neighbourhood (1 for the most frequent and so forth). Figure \ref{fig:dictionary} shows some frequent neighbourhoods that will be discussed later.
% Computational cost
In the worst case, where every post has a unique neighbourhood, every extracted neighbourhood should be compared against every neighbourhood that already appears in the dictionary before concluding that it needs a new entry. In this case, the computational cost of this operation is $\mathcal{O}(n^2)$ where $n$ is the number of posts. In practice, however, the bigger the number of posts processed, the less likely we go through all the dictionary without finding an isomorphism. Moreover, if we keep the dictionary sorted by frequency, a large number of neighbourhoods will find an isomorphism among the first neighbours of the dictionary.

@@

\begin{table}
 \begin{center}
   \tabcolsep = 0.8\tabcolsep
   \begin{tabular}{|rr|rr|rr|}
      \hline
      \multicolumn{2}{|c|}{struct-based} & \multicolumn{2}{|c|}{order-based} &  \multicolumn{2}{|c|}{time-based} \\
      %\hline
     	 POD & GOT 						& POD & GOT 		& POD & GOT\\
   \hline \hline
     7 (9.6)	& 7 (7.4)   		&  		1 (25.2) & 1 (25.2) &  4 (9.0) & 4 (6.4)\\    
     3 (16.5)& 9 (12.6)   	&  	3 (32.1) & 3 (32.1) &  8 (17.2) & 8 (12.2)\\ 
     2 (23.1)& 3 (17.8)   	&  	2 (38.8) & 2 (38.8) &  2 (25.3) & 2 (17.9)\\ 
     20 (27.8)& 11 (22.9)   &  	13 (45.3) & 13 (45.3) &  3 (31.2) & 9 (23.2)\\    
     4 (31.1)& 5 (27.8)   	&  	4 (50.1) & 4 (50.1) &  10 (36.6) & 14(28.4)\\ 
     13 (34.1)& 2 (32.3)   	&  	18 (54.7) & 18 (54.7) &  14 (41.8) & 11(33.5)\\ 
     6 (36.9)& 12 (36.1)   	&  	6 (58.9) & 6 (58.9) &  6 (45.8) & 10(38.5)\\    
     5 (39.4)& 27 (38.8)   	&	  	12 (65.5) & 10 (62.8) &  15 (48.9) & 5 (43.1)\\ 
     15 (41.9)& 4 (41.4)   	&  	16 (68.5) & 12 (65.5) &  19 (51.5) & 6 (47.4)\\  
     12 (43.7)& 6 (43.9)   	&  	5 (70.6) & 16 (68.1) &  5 (53.9) & 3 (51.3)\\    
     21 (45.3)& 21 (45.9)   	&  	17 (72.8) & 5 (70.7) &  7 (55.9) & 25(55.1)\\ 
     30 (46.9)& 15 (47.7)   	&  	23 (75.5) & 17 (72.8)&  16 (57.6) & 15(57.8)\\ 
     16 (48.4)& 13 (49.2)   	&  	31 (76.1) & 23 (74.5)&  36 (59.4) & 12(60.2)\\    
     39 (49.8)& 20 (50.6)   	&  	19 (77.5) & 31 (76.1) &  25 (61.2) & 7 (62.2)\\ 
     44 (51.2)& 30 (51.9)   	&  	22 (79.0) & 19 (77.5) &  21 (62.6) & 21(63.9)\\ 
     8 (52.4)& 22 (53.1)   	&	  	24 (80.3) & 22 (78.9) &  34 (63.9) & 19(65.5)\\    
     22 (53.3)& 43 (54.2)   	&  	26 (81.5) & 24 (80.3) &  38 (65.1) & 17(66.7)\\ 
     51 (54.2)& 8 (55.2)   	&  	29 (82.7)& 26 (81.5) &  23 (66.4) & 36(68.0)\\          
   \hline
   \end{tabular}
\label{tab:table}
 \end{center}
 \caption{Frequence this is an example}
\end{table}
